---
title: "Group 6 Assignment 1: R Programming"
output: 
  pdf_document: 
    toc: yes
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo = F}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```


\newpage

## R Markdown

```{r echo=TRUE}
library(ggpubr)
library(tidyverse)
library(janitor) #janitor helps us clean datasets
library(here) # here helps us know where files are
library(readxl)
library(readr)
library('formatR')
```


## Load data - CSV
## This data is a COVID dataset and downloaded open data
```{r}
COVID19_Dataset <- readr::read_csv("https://opendata.ecdc.europa.eu/covid19/nationalcasedeath/csv")
```

## Print the structure of your dataset
```{r}
str(COVID19_Dataset)
```
## List the variables in your dataset
```{r}
names(COVID19_Dataset)
```
## Print the top 15 rows of your dataset
```{r}
head(COVID19_Dataset, n = 15)
```
## Write a user defined function using any of the variables from the data set
```{r}
my_function <- function(x, y) {
output <- COVID19_Dataset %>%  group_by(indicator, continent, year_week) %>%
  summarize(mean_weekly_count = mean(weekly_count, na.rm = TRUE)) %>%
  filter(indicator == x, year_week == y)
return(output)
}

my_function('deaths', '2021-01')

my_function('cases', '2020-12')
```

## Use data manipulation techniques and filter rows based on any logical criteria that exist in your dataset
```{r}
COVID19_Dataset_Canada = COVID19_Dataset %>% filter(COVID19_Dataset$weekly_count >= 1000, COVID19_Dataset$country == "Canada")

head(COVID19_Dataset_Canada,n=10)
```
## Identify the dependent & independent variables and use reshaping techniques and create a new data frame by joining those variables from your dataset
```{r}
indept = data.frame(COVID19_Dataset$population, COVID19_Dataset$country, COVID19_Dataset$year_week)
dept = data.frame(COVID19_Dataset$weekly_count)
new_set = cbind.data.frame(indept, dept)
head(new_set, n=10)
```
## Remove missing values in your dataset
```{r}
COVID19_Dataset_notnull = COVID19_Dataset[complete.cases(COVID19_Dataset), ]
head(COVID19_Dataset_notnull, n=10)
```
## Identify and remove duplicated data in your dataset
```{r}
COVID19_Dataset_distinct = COVID19_Dataset %>% distinct() 
head(COVID19_Dataset_distinct, n=10)
```
## Reorder multiple rows in descending order
```{r}
head(COVID19_Dataset %>% arrange(desc(COVID19_Dataset$country, COVID19_Dataset$continent)),n=10)
```
## Rename some of the column names in your dataset
```{r}
COVID19_Dataset_renamed = COVID19_Dataset %>% rename( Country_Name = country, Cotinent_Name=continent)
head(COVID19_Dataset_renamed,n=10)
```
## Add new variables in your data frame by using a mathematical function (for e.g. – multiply an existing column by 2 and add it as a new variable to your data frame)
```{r}
COVID19_Dataset_added = COVID19_Dataset %>% mutate(count_by_population = cumulative_count/population)
head(COVID19_Dataset_added,n=10)
```
## Create a training set using random number generator engine
```{r}
set.seed(1234)
trainingset = COVID19_Dataset %>% sample_frac(0.05, replace = FALSE)
head(trainingset,n=10)
```

## Print the summary statistics of your dataset
```{r}
COVID19_Dataset %>% group_by(COVID19_Dataset$continent) %>% summarise_if(is.numeric, median, na.rm = TRUE)
```

## Use any of the numerical variables from the dataset and perform the following statistical functions. 
## Mean • Median • Mode • Range
```{r echo=TRUE}
mode(COVID19_Dataset$cumulative_count)
range(COVID19_Dataset$cumulative_count)
median(COVID19_Dataset$cumulative_count)
mean(COVID19_Dataset$cumulative_count)
```
## Plot a scatter plot for any 2 variables in your dataset.
```{r}

my_function2 <- function(x) {
output <- COVID19_Dataset %>%  group_by(indicator, continent, year_week) %>%
  summarize(mean_weekly_count = mean(weekly_count, na.rm = TRUE)) %>%
  filter(indicator == x)
return(output)
}

Plot_data = my_function2('deaths')


ggplot(data = Plot_data,aes(x = year_week, y = mean_weekly_count, 
                      col = factor(Plot_data$continent)))+geom_point()

```
## Plot a bar plot for any 2 variables in your dataset
```{r}

plot_data2 = Plot_data %>% filter(year_week == "2020-12")
ggplot(data = plot_data2, aes(continent, mean_weekly_count)) + geom_bar(stat = "identity", fill = "aquamarine4")
```



## Find the correlation between any 2 variables by applying least square linear regression model
```{r echo=TRUE, message=FALSE, out.width='50%', tidy=TRUE}

COVID19_Dataset_cases = 
  COVID19_Dataset %>% filter(indicator == "cases", 
                             year_week == '2020-12'|year_week =='2020-13'|year_week =='2020-14'|year_week == '2020-15'|year_week =='2020-16'|year_week =='2020-17'|year_week =='2020-18'|year_week =='2020-19'|year_week =='2020-20')

COVID19_Dataset_cases_renamed = COVID19_Dataset_cases %>% rename(  indicator_count_cases = weekly_count)

COVID19_Dataset_deaths = 
  COVID19_Dataset %>% filter(indicator == "deaths", 
                       year_week == '2020-12'|year_week =='2020-13'|year_week =='2020-14'|year_week == '2020-15'|year_week =='2020-16'|year_week =='2020-17'|year_week =='2020-18'|year_week =='2020-19'|year_week =='2020-20')

COVID19_Dataset_deaths_renamed = COVID19_Dataset_deaths %>% rename(  indicator_count_deaths = weekly_count)

new_data_set = cbind.data.frame(COVID19_Dataset_cases_renamed, COVID19_Dataset_deaths_renamed$indicator_count_deaths)

X = new_data_set[,"indicator_count_cases"]
Y = new_data_set[,"COVID19_Dataset_deaths_renamed$indicator_count_deaths"]
correlation = cor(Y,X,method = "pearson")
ggscatter(new_data_set, x = "indicator_count_cases", 
          y = "COVID19_Dataset_deaths_renamed$indicator_count_deaths", 
add = "reg.line",conf.int = TRUE, cor.coef = TRUE, cor.method = "pearson", 
xlab = "indicator_count_cases", ylab = "indicator_count_deaths")


```
```

## Provide a conclusion of your analysis if any in the .RMD file
```{r}
#According to our scatter plot, different continents have given different responses to Covid-19. For example, Oceania has the lowest death toll. On the flip side, Europe has completely different counts. It can easily be seen that, in the middle of Covid-19, Europe took some precautions, and decreased the deaths toll. Additionally, we can say that America has the highest death toll.

#When we look at the correlation between cases and deaths, we can say that two variable depend on each other. They have a high correlation according to the correlation graphic.At the 12 th week of 2020, every case almost ended up with a death. But after a while, 20 th week of 2021, the death toll is lower than case counts. We can say that during these times, people has increased their immune system.
```



